\abstract{
Most Internet censorship research measures the network layer: whether a site is
DNS-poisoned, IP-blocked, or throttled. Yet a growing share of censorship now
occurs above the network, at the mobile app stores operated by Apple and Google.
Because both companies run a separate storefront for every country, a government
that wants to make an app disappear no longer needs to filter traffic; it can
simply compel the platform to remove the app from its national store. We report a
snapshot measurement of this ``new chokepoint.'' We built a pipeline that queries
the per-country storefronts of the Apple App Store for a curated basket of 32
apps (VPNs, secure messengers, news, topically sensitive apps, and neutral
controls) across 14 countries, and cross-validated the results against documented
removal events. Our snapshot independently reproduces every documented removal we
tested, including the 2024 Russian VPN purge, the 2017 China VPN and New York
Times removals, and Signal's absence from the China store. It also surfaces a
clean cross-country signal: sensitive-app unavailability is negatively correlated
with Freedom House's \emph{Freedom on the Net} score (Spearman $r=-0.62$), driven
by near-total removal in China and a VPN-focused purge in Russia. A central
methodological contribution is an explicit treatment of the attribution problem:
distinguishing a government-compelled removal from a sanctions-driven developer
withdrawal or a benign regional app variant. We release an open dataset and
reproducible method that complements network-level tools such as OONI at a layer
they do not currently cover.
}
