\section{Project Preparation and Prerequisites}\label{sec:approach}

This is a measurement-based project. Below we describe the data we already have
access to, the data we will collect, the tooling we will build or extend, and the
results we expect to produce and how we will present them.

\subsection{Data We Have}

The core measurement surface is public and queryable from any vantage point,
which is a significant practical advantage: we do not need machines or volunteers
inside censored countries. Specifically:
\begin{itemize}
  \item \textbf{Apple App Store.} The public iTunes Search and Lookup API accepts
  a \texttt{country} parameter, so the same app identifier can be queried against
  every national storefront and will return a result where the app is available
  and an empty result where it is not.
  \item \textbf{Google Play.} Google exposes no equally clean API, but mature
  open-source scrapers (e.g., \texttt{google-play-scraper}) and the prior
  \texttt{appavail} tooling released with Ververis et
  al.~\cite{ververis2019shedding} allow per-country availability checks.
  \item \textbf{Ground-truth references.} Apple and Google publish transparency
  reports listing government-driven takedown requests; GreatFire's
  AppleCensorship provides historical removal records~\cite{applecensorship}; and
  Freedom House's \emph{Freedom on the Net} provides per-country repression
  scores we can correlate against.
\end{itemize}

\subsection{Data We Will Collect}

We will assemble two fixed lists and cross them:
\begin{itemize}
  \item An \textbf{app basket} of roughly 40--60 apps spanning categories that
  governments are known to target: VPNs (e.g., NordVPN, Proton VPN, Tor), secure
  messengers (Signal, Telegram, WhatsApp), independent and foreign news (NYT,
  BBC, regional outlets), and topically sensitive apps (LGBTQ+, Tibetan, and
  religious apps). Critically, we include a set of \textbf{neutral control apps}
  (e.g., a calculator, a weather app) that no government is expected to censor.
  \item A \textbf{country list} covering all storefronts we can query, with
  emphasis on known censors (China, Russia, Iran, Turkey, UAE, India) and
  free-expression baselines (US, Germany, Canada).
\end{itemize}
Querying the full cross product yields an availability matrix (apps $\times$
countries). We will re-run the collection \textbf{weekly over the term} so that we
capture removals and reinstatements as time series rather than a single snapshot.

\subsection{Tooling}

Rather than build from scratch, we will extend the \texttt{appavail} scripts from
prior work~\cite{ververis2019shedding} and the \texttt{google-play-scraper}
library, wrapping them in a scheduler that snapshots results, diffs consecutive
runs, and stores everything in a versioned dataset. Analysis will use Python and
Pandas. All code and the resulting dataset will be released openly.

\subsection{Expected Results and Presentation}

We expect to produce: (1) a \textbf{country $\times$ app availability heatmap},
where censorship appears as dark bands over particular countries or categories;
(2) a \textbf{timeline} of any removals or reinstatements observed during the
term; and (3) a \textbf{correlation analysis} between per-country missing-app
counts and independent repression indices such as \emph{Freedom on the Net}. Even
a ``null'' result (a clean, validated map of what is available where) is a
contribution, because no one currently maintains such a dataset systematically
across categories.

\subsection{Prerequisites and Feasibility}

The project requires only Python/Pandas proficiency and familiarity with HTTP
APIs and light scraping, all within the team's background. There is no dependency
on hardware, field partners, or third parties, so the timeline is under our
control: scraper construction in weeks 1--2, collection over the following weeks,
and analysis and writing at the end. The main technical risks are API rate limits
and undocumented changes to store endpoints, both of which we mitigate by
throttling requests, caching raw responses, and cross-checking two independent
data sources (direct queries and AppleCensorship records).
